Supplimentary figures

### 2020-08-14 manuscript.Rmd
for mac

authornote: |
    12961 words in text body and refence section

    Number of tables: 11

    Number of figures: 10

@@

Over the last four decades sport-related injuries have received increased research attention [@Ivarsson2017] in response to the high incidence [@Rosa2014; @Sheu2016] and associated undesirable physical and psychological effects [@Leddy1994; @Brewer2012]. Multiple psychological [@Slimani2018], anatomical [@Murphy2003], biomechanical [@Neely1998; @Hughes2014] and environmental [@Meeuwisse2007] factors have been associated with sports injury occurrence and several models of injury causation have been proposed that highlight the multifactorial nature of injury occurrence [@Kumar2001; @Meeuwisse2007; @Wiese-Bjornstal2009], of which one of the most widely cited was developed by Williams and Anderson [Fig \ref{fig:fig1}; @Andersen1988; @Williams1998].

```{r fig1, fig.cap="Stress and injury model (Williams and Andersen, 1998).", fig.align='center',fig.pos = "!h"}
include_graphics("figures_doc/Fig1.jpeg")
```

Williams and Andersen’s [@Williams1998] stress-injury model proposed that when faced with a potentially stressful athletic situation, an athlete’s personality traits (e.g., hardiness, locus of control and competitive trait anxiety), history of stressors (e.g., major life events and previous injuries) and coping resources (e.g., general coping behaviours) contribute to the injury response, either interactively or in isolation. The stress response is central to the model and reflects the bi-directional relationship between athletes’ appraisal of, and response to, a stressful athletic situation. The model predicts that athletes who have a history of stressors, personality traits that intensify the stress response and few coping resources, will exhibit greater attentional (e.g., peripheral narrowing) and/or physiological (e.g., increased muscle tension) responses that put these individuals at greater injury risk.

@@

Postural stability was assessed with a modified version of the balance error scoring system (mBESS) based on the protocol recommended by @Hunt2009. 
In total, each trial of the mBESS was performed without shoes [@McCrory2013] and included six stances in the following order; dominant leg (DL; standing on the dominant foot with the non-dominant foot at approximately 30-degrees of hip flexion and 45-degrees of knee flexion),
non-dominant leg (NDL; standing on the non-dominant foot with the dominant foot at approximately 30-degrees of hip flexion and 45-degrees of knee flexion) 
and tandem leg stance (TS; standing heel-to-toe with the non-dominant foot behind the dominant) on firm and foam (Alcan airex AG, Sins, Switzerland) surfaces respectively (Fig \ref{fig:fig2}). 
To determine leg dominance, participants were asked their preferred leg to kick a ball to a target, and the chosen limb was labelled as dominant [cf. @VanCingel2017]. 
Participants were asked to maintain each stance for a total of 20-seconds.
Participants hands were placed on hips at the level of the iliac crests. 
A Sony DSC-RX10 video camera (Sony Europe Limited, Surrey, United Kingdom) was used to record each participants performance during the mBESS.

```{r fig2, fig.cap="mBESS positions (A-F). Top row, firm surface. Bottom row, foam surface. Left column, dominant leg stance. Middle column, non-dominant leg stance. Right column, Tandem leg stance.", out.width = "100%", fig.align='center',fig.pos = "!h"}
include_graphics("figures_doc/Fig2.jpeg")
```

@@

At the start of the academic year (September), coaches of sports teams at a British University and local sports clubs were contacted and informed about the study. With the coaches’ permission, the lead researcher attended training sessions to inform athletes about the overall purpose of the study and the requirements of participation. Athletes who met the participation criteria and volunteered to take part in the study were invited to attend scheduled testing sessions. A repeated measures prospective cohort design was used to assess athletes’ major life events, stress-related physiological markers and injury status over two consecutive 12-month periods. Each participant was asked to attend a total of three data collections over a 12-month period, with each data collection separated by a four-month interval (Fig \ref{fig:fig3}). Participants provided informed consent before data collection commenced.

```{r fig3, fig.cap = 'Study design. For each time point (T), each box contains the number of participants who completed the data collection (\\textit{n}), the measures used for data collection and the approximate date of the data collection.', out.width = "100%"}

include_graphics("figures_doc/Fig3.jpeg")
```

For the first three data collections (T1, T2 and T3), participants followed the same protocol in a specific order (Fig \ref{fig:fig4}). 
To ensure all measures could be collected within a viable time-frame, participants were separated into two groups. 
The first group completed all computer-based measures followed by all physical measurements, whereas the second group completed all physical measurements followed by computer-based measures. 
Participants were randomly assigned to one of the two groups and remained in those groups across all time points.

```{r fig4, fig.cap = 'Outline of the protocol for each data collection.', out.width="100%", fig.align='center'}
include_graphics("figures_doc/Fig4.jpeg")
```

@@

Learning the structure of the network is an important step in BN modelling. The structure of a network can be constructed using expert knowledge and/or data-driven algorithm techniques (e.g., search and score, such as hill climbing and gradient descent algorithms; Scutari & Denis, 2014).
The learned structure can then be used for inference by querying the network[^1] and obtaining the posterior probabilities of a particular node for a given query.
The posterior distribution can be obtained by $Pr(X|E,B) = Pr(X|E,G,\Theta)$, where the learned network $B$ with structure $G$ and parameters $\Theta$, are investigated with new evidence $E$ using the information in $B$ [@Scutari2014]. 
In the example network presented in Fig \ref{fig:fig5}, new values assigned to each of the parent nodes (e.g., both set to "Low") could be used to investigate what effect the new information has on the state of the child node (conditional probability of a particular state of the child node).
In a more complex network containing many nodes, the outcome of a particular node can be assessed conditional on the states of any subset of nodes in the network. 
BNs therefore provide a unique and versatile approach to modelling a set of variables to uncover dependency structures within the data.

@@

The subscales representing the BAS (RR, RI, GDP and I) showed limited connection to other variables in the network. 
Therefore, competing models were examined and BIC scores compared to establish the model with the best fit to the data (values are shown relative to the highest value).
RI provided the highest BIC value (`r rstmodel[1,2]`), compared to RR (`r rstmodel[2,2]`), GDP (`r rstmodel[4,2]`) and I (`r rstmodel[5,2]`).
Including all the variables (RR, RI, GDP and I) resulted in a significantly lower score `r rstmodel[6,2]`) indicating that including all the variables was not beneficial to the model structure and did not offset the cost of the additional parameters. Therefore, only RI was included in the final structure. 
Finally, both total score and asymmetry for balance were included in the initial network. However, visual inspection of the network revealed no arcs between bal_asym_1 or bal_asym_2 and any other node in the network.
Therefore, balance asymmetry was removed from the final network structure.
To summarise, Table \ref{tab:table3} reports the variables that were included in the final network structure.

```{r table3, results='asis'}
varcutoffs %>% 
  filter(!var %in% c("sdnn", "RR", "I", "GDP", "tlelg", "tlelg_1", "tlelg_2", "tlelg_3", "bal_asym")
) %>% 
  mutate(var = ifelse(var == "hrv", "HRV", var)) %>% 
  kable("latex",
       align = c("llcc"),
       caption = "Variables included in the final Bayesian network structure.",
       col.names = c("Variable", "Definition", "State 1", "State 2")) %>%
  kable_styling(latex_options = "HOLD_position") %>% 
  column_spec(2, width = c("5.2cm")) 
```

@@

During the study, 46% (_n_ = 117) of participants reported at least one injury with an average severity of `r mean_sd(datanet$days_missed, digits = 0)`, days (range = `r min(datanet$days_missed[datanet$days_missed > 1])` - `r max(datanet$days_missed)` days).
Both male and female participants reported a greater number of acute compared to chronic injuries (male, acute = `r inj_mechanism[1,3]`, chronic = `r inj_mechanism[2,3]`; female, acute = `r inj_mechanism[1,2]` chronic = `r inj_mechanism[2,2]`), and non-contact injuries were more common than contact injuries (male, non-contact = `r inj_mechanism[4,3]`, contact = `r inj_mechanism[2,3]`; female, non-contact = `r inj_mechanism[4,2]` contact = `r inj_mechanism[3,2]`).
Table \ref{tab:table4} shows the number and percentage of injury types sustained by both male and female participants.

@@

```{r table4, results='asis'}

@@

The first network structure obtained from the data (Fig \ref{fig:fig6}) examined the interactions between explanatory variables, independent variables and probability of injury across time points in the study. 
Strong sport-related and gender-based connections between several explanatory and independent variables were demonstrated for individual and team-based sports.
The "ind_team" node had strong arcs to hours (`r str_finder("ind_team", "hours")`) and nlebase (`r str_finder("ind_team", "nlebase")`).
Individual athletes were more likely to have "High" hours per week (`r x[5,1]`) compared to team-based athletes (`r x[6,1]`).
Individual athletes were also more likely to have "High" negative life events in the 12 months preceding the start of the study compared to team-based athletes (individual athletes = `r x[7,1]`, team-based athletes = `r x[8,1]`).
The arc from clevel $\rightarrow$ balance_1 had a strength of `r str_finder("clevel", "balance_1")`, with lower level performers more likely to have decreased balance ability (`r x[21,1]`), compared to national level athletes (`r x[22,1]`). 
High gender-based connections were reported for the arcs from gender  $\rightarrow$ stiffness_1 (`r str_finder("gender", "stiffness_1")`) and gender $\rightarrow$ stiffness_2 (`r str_finder("gender", "stiffness_2")`), with males more likely to have "High" stiffness compared to females (males = `r x[14,1]`, females = `r x[13,1]`).
Irrespective of sport or gender, strong connections were found between explanatory variables. The arc from nlebase $\rightarrow$ RI_1 had a strength of `r str_finder("nlebase", "RI_1")`, and the probability of RI_1 being in the “High” state increased from `r x[2,1]` to `r x[3,1]` when nlebase increased from “Low” to “High”.

@@

```{r fig6, fig.align='default', fig.height=10, fig.width=15, fig.cap='The full Bayesian network structure was plotted using the _strength.plot_ function in bnlearn. The strength of each arc is shown graphically by the style of the arc. Thin, dashed arcs indicate the weakest arcs (arc strength below 0.50), whereas thick solid arcs indicate the strongest arcs (arc strength of 1). White nodes in the network indicate the explanatory variables, blue nodes indicate T1_1 and T2_1 variables, and red nodes indicated T2_2 and T3_2 variables. The injured_X nodes have been coloured gold as they are the main nodes of interest within the network.'}
renderGraph(gR)
```

@@

The first conditional probability query (CPQ) investigated the variables in the Markov blanket for injured\_1 (Fig \ref{fig:fig7}), which contained hours spent training per week (hours), negative life events (nlelg_1), muscle stiffness (stiffness_1), current competitive level (clevel) and balance (balance_1). 
The arc between nlelg_1 and injured_1 was fixed in the network, so has the maximum strength of 1.

```{r fig7, fig.cap="Markov blanket of injured\\_1. Arc strengths are included as arc labels.", out.width="100%", fig.align="center"}
include_graphics("figures_doc/Fig7.jpeg")
```

The CPQ for injured_1 in the "injured" state for all variables that were linked to injured_1 is shown in Table \ref{tab:table5}.
The probability of injured_1 = "injured" rose from `r query1[4,2]` to `r query1[4,3]` when stiffness was "High" compared to "Low".
Negative life events had a negligible effect when moving from the "Low" to "High" state.

```{r table5, results='asis'}
query1 %>%
  select(c(1, 2, 3)) %>%
  mutate(variable = str_replace_all(variable, "_", "\\\\_")) %>%
  kable(
    .,
    caption = "Probability of injured\\_1 being in the ``injured'' state, conditional on each variable",
    col.names = c("Variable", "Low", "High"),
    escape = F
  ) %>%
  kable_styling(latex_options = "HOLD_position") %>%
  row_spec(0,bold=TRUE)
```

The second CPQ investigated the outcome of injured_1 being "injured" conditional on all variables in the Markov blanket.
The Markov blanket contained five nodes, each with two possible states resulting in $2^5$ combinations of variables, therefore only the three lowest and highest probabilities are shown in Table \ref{tab:table6} (complete results in supplementary table 4).
The combination of lower competitive level, "High" hours per week, "Low" negative life events, "High" balance and "High" stiffness resulted in a probability of `r max(query2$prob)` for injured_1 being in the "injured" state.
When all variables were in the "Low" state the probability of "injured" was approximately `r min(query2$prob)`

@@

```{r table6, results='asis'}

query2 %>%
  select(7, 2:6) %>%
  rename(Probability = "prob") %>%
  `colnames<-`(str_replace_all(colnames(.), "_", "\\\\_")) %>%
  mutate(clevel = str_replace_all(clevel, "_", "\\\\_")) %>%
  kable(
    .,
    caption = "Highest and lowest probability of injured\\_1 being in the 'injured' state, conditional on the all variables in the Markov blanket for injured\\_1.",
    escape = F,
    align = "ccccc"
  ) %>%
  kable_styling(latex_options = "HOLD_position") %>%
  pack_rows("Highest", 1, 3) %>%
  pack_rows("Lowest", 4, 6) %>%
  row_spec(0,bold=TRUE)
```

@@

The Markov blanket for injured_2 is shown in Figure \ref{fig:fig8} and contained gender, FFFS_1, stiffness_2, balance_2 and hrv_2.
The arc between stiffness_2 $\rightarrow$ injured_2 was comparable to the arc between stiffness_1 $\rightarrow$ injured_1.
Very weak arcs (0.3) between injured_2 $\rightarrow$ balance_2 and injured_2 $\rightarrow$ hrv_2 were also present in the Markov blanket for injured_2.
Similar to injured_1, stiffness_2 doubled the probability of injured_2 being "injured" from `r query3[5,2]` in the "Low" state to `r query3[5,3]` in the "High" state.
FFFS_1 in the "Low" state increased probability of injured_2 being "injured" by `r query3[2,2]-query3[2,3]` compared to the "High" state.
"High" negative life events decreased the probability of injury from `r query3[3,2]` to `r query3[3,3]`.

```{r table7, results="asis"}

query3 %>%
  mutate(Low = format(Low, nsmall = 2)) %>% 
  mutate_all(~str_replace_all(., "_", "\\\\_"))%>%
  kable(
    .,
    caption = "Probability of injured\\_2 being in the ``injured' state, conditional on each variable in the Markov blanket for injured\\_2.",
    col.names = c("Variable", "Low", "High"),
    escape = F,
    align = "cccccc"
  ) %>%
  kable_styling(latex_options = "HOLD_position") %>%
  row_spec(0,bold=TRUE)
```

```{r fig8, fig.cap="Markov blanket for injured\\_2.", out.width="100%", fig.align="center"}
include_graphics("figures_doc/Fig8.jpeg")
```

@@

The three lowest and highest conditional probabilities based on all the variables in injured_2 Markov blanket are presented in Table \ref{tab:table8} (complete results in supplementary table 5).
The combination of "Low" FFFS_1, "High" stiffness_2, "High" balance resulted in the greatest  probability of injured_2 being "injured", with the highest probability of injury being `r max(query4$prob)`.
With all other variables held in the "High" state, the probability of injured_2 being "injured" rose from `r min(fffs_query$prob)` to `r max(fffs_query$prob)` when FFFS_1 was in the "Low" compared to "High" state.
The combination of "Low" stiffness, "Low" balance and "High" FFFS resulted in the lowest probability of injured_2 being "injured".

```{r table8, results="asis" }

query4 %>% 
  toptail(3) %>%
  select(7,2:6) %>%
  rename(Probability = "prob") %>%
  `colnames<-`(str_replace_all(colnames(.), "_", "\\\\_")) %>%
  mutate_all(~str_replace_all(., "_", "\\\\_"))%>%
  kable(
    .,
    caption = "Highest and lowest probability of injured\\_2 being in the 'injured' state, conditional on the all variables in the Markov blanket for injured\\_2.",
    escape = F,
    align = "cccccc"
  ) %>%
  kable_styling(latex_options = "HOLD_position") %>%
  pack_rows("Highest", 1, 3) %>%
  pack_rows("Lowest", 4, 6) %>%
  row_spec(0,bold=TRUE)
```

@@

The second network structure (Fig \ref{fig:fig9}) examined changes within variables between time points and the probability of injury/
An arc between BIS $\rightarrow$ FFFS with strength `r str_cfinder("BIS", "FFFS")` was present in the network.
Arcs between clevel $\rightarrow$ BIS and gender $\rightarrow$ stiffness had a strengths of `r str_cfinder("clevel", "BIS")` and `r str_cfinder("gender", "stiffness")` respectively
Similarly, the arc between hrv $\rightarrow$ FFFS was `r str_cfinder("hrv","FFFS")`.
The arcs between BIS $\rightarrow$ FFFS and hrv $\rightarrow$ FFFS were examined further by drawing random observations from the conditional probability distribution and examining the relationship in a Bayesian linear regression model.
A separate linear regression examined the interaction between BIS and hrv to be examined.

```{r fig9, fig.cap="Network structure of the changes within variables between time points.", out.width="100%", fig.align="center"}

include_graphics("figures_doc/Fig9.jpeg")

@@

Results from the Bayesian linear regression model are presented in Table \ref{tab:table9} and include 95% credible intervals (CrI).
Increases in BIS were associated with increases in FFFS ($b$ = `r bmtable[3,2]`, 95% CrI = `r bmtable[3,4]`), whereas positive changes in hrv where associated with decreased changes in FFFS ($b$ = `r bmtable[2,2]`, 95% CrI = `r bmtable[2,4]`).
There was no clear interaction between hrv and BIS ($b$ = `r bmtable[4,2]`, 95% CrI = `r bmtable[4,4]`).

```{r table9, results="asis"}
bmtable %>%
  kable(
    .,
    caption = "Estimate, error and 95\\% credible intervals for the fixed effects in the linear model containing FFFS, BIS and hrv.",
    col.names = c("Term", "Estimate", "Error", "95\\% CI"),
    escape = F,
    align = "lccc"
  ) %>%
  kable_styling(latex_options = "HOLD_position") %>%
  row_spec(0,bold=TRUE)

@@

The Markov blanket for the "injured" node contained previous injury, gender, training hours per week, stiffness and nlec (Fig \ref{fig:fig10}).
For stiffness and nlec the values in the nodes represent the standardised change between time point.
Combinations of nlec and stiffness at one SD below the mean change, at the mean change, and 1 SD above the mean change are presented in table \ref{tab:table10}.
Increases in muscle stiffness was found to increase the risk of injury, which was further increased when there were increases in NLE stress. 
Changes in both nlec and stiffness of 1SD above the mean change resulted in a high probability of being injured (`r query6a[9,1]`) over the duration of the study.
With stiffness held at the mean change, the probability of "injured" rose notably from `r query6a[4,1]` to `r query6a[6,1]` with nlec at 1 SD below an 1 SD above respectively.

```{r fig10, fig.cap="Markov blanket for the injured node in the network reflecting changes within variables between time points.", out.width="100%", fig.align="center"}
include_graphics("figures_doc/Fig10.jpeg")
```

>>>>

authornote: |
    12856 words in text body and refence section

    Number of tables: 10

    Number of figures: 5

@@

Over the last four decades sport-related injuries have received increased research attention [@Ivarsson2017] in response to the high incidence [@Rosa2014; @Sheu2016] and associated undesirable physical and psychological effects [@Leddy1994; @Brewer2012]. Multiple psychological [@Slimani2018], anatomical [@Murphy2003], biomechanical [@Neely1998; @Hughes2014] and environmental [@Meeuwisse2007] factors have been associated with sports injury occurrence and several models of injury causation have been proposed that highlight the multifactorial nature of injury occurrence [@Kumar2001; @Meeuwisse2007; @Wiese-Bjornstal2009], of which one of the most widely cited was developed by Williams and Anderson [@Andersen1988; @Williams1998].

Williams and Andersen’s [@Williams1998] stress-injury model proposed that when faced with a potentially stressful athletic situation, an athlete’s personality traits (e.g., hardiness, locus of control and competitive trait anxiety), history of stressors (e.g., major life events and previous injuries) and coping resources (e.g., general coping behaviours) contribute to the injury response, either interactively or in isolation. The stress response is central to the model and reflects the bi-directional relationship between athletes’ appraisal of, and response to, a stressful athletic situation. The model predicts that athletes who have a history of stressors, personality traits that intensify the stress response and few coping resources, will exhibit greater attentional (e.g., peripheral narrowing) and/or physiological (e.g., increased muscle tension) responses that put these individuals at greater injury risk (supplementary figure 1).

@@

Postural stability was assessed with a modified version of the balance error scoring system (mBESS) based on the protocol recommended by @Hunt2009. 
In total, each trial of the mBESS was performed without shoes [@McCrory2013] and included six stances in the following order; dominant leg (DL; standing on the dominant foot with the non-dominant foot at approximately 30-degrees of hip flexion and 45-degrees of knee flexion),
non-dominant leg (NDL; standing on the non-dominant foot with the dominant foot at approximately 30-degrees of hip flexion and 45-degrees of knee flexion) 
and tandem leg stance (TS; standing heel-to-toe with the non-dominant foot behind the dominant) on firm and foam (Alcan airex AG, Sins, Switzerland) surfaces respectively (supplementary figure 2). 
To determine leg dominance, participants were asked their preferred leg to kick a ball to a target, and the chosen limb was labelled as dominant [cf. @VanCingel2017]. 
Participants were asked to maintain each stance for a total of 20-seconds.
Participants hands were placed on hips at the level of the iliac crests. 
A Sony DSC-RX10 video camera (Sony Europe Limited, Surrey, United Kingdom) was used to record each participants performance during the mBESS.

@@

At the start of the academic year (September), coaches of sports teams at a British University and local sports clubs were contacted and informed about the study. With the coaches’ permission, the lead researcher attended training sessions to inform athletes about the overall purpose of the study and the requirements of participation. Athletes who met the participation criteria and volunteered to take part in the study were invited to attend scheduled testing sessions. A repeated measures prospective cohort design was used to assess athletes’ major life events, stress-related physiological markers and injury status over two consecutive 12-month periods. Each participant was asked to attend a total of three data collections over a 12-month period, with each data collection separated by a four-month interval (supplementary figure 3). Participants provided informed consent before data collection commenced.

For the first three data collections (T1, T2 and T3), participants followed the same protocol in a specific order (supplementary figure 4). 
To ensure all measures could be collected within a viable time-frame, participants were separated into two groups. 
The first group completed all computer-based measures followed by all physical measurements, whereas the second group completed all physical measurements followed by computer-based measures. 
Participants were randomly assigned to one of the two groups and remained in those groups across all time points.

@@

Learning the structure of the network is an important step in BN modelling. The structure of a network can be constructed using expert knowledge and/or data-driven algorithm techniques (e.g., search and score, such as hill climbing and gradient descent algorithms; Scutari & Denis, 2014).
The learned structure can then be used for inference by querying the network[^1] and obtaining the posterior probabilities of a particular node for a given query.
The posterior distribution can be obtained by $Pr(X|E,B) = Pr(X|E,G,\Theta)$, where the learned network $B$ with structure $G$ and parameters $\Theta$, are investigated with new evidence $E$ using the information in $B$ [@Scutari2014]. 
When a network contains many nodes, the outcome of a particular node can be assessed conditional on the states of any subset of nodes in the network. 
BNs therefore provide a unique and versatile approach to modelling a set of variables to uncover dependency structures within the data.

@@

The subscales representing the BAS (RR, RI, GDP and I) showed limited connection to other variables in the network. 
Therefore, competing models were examined and BIC scores compared to establish the model with the best fit to the data (values are shown relative to the highest value).
RI provided the highest BIC value (`r rstmodel[1,2]`), compared to RR (`r rstmodel[2,2]`), GDP (`r rstmodel[4,2]`) and I (`r rstmodel[5,2]`).
Including all the variables (RR, RI, GDP and I) resulted in a significantly lower score `r rstmodel[6,2]`) indicating that including all the variables was not beneficial to the model structure and did not offset the cost of the additional parameters. Therefore, only RI was included in the final structure. 
Finally, both total score and asymmetry for balance were included in the initial network. However, visual inspection of the network revealed no arcs between bal_asym_1 or bal_asym_2 and any other node in the network.
Therefore, balance asymmetry was removed from the final network structure.
To summarise, Table \ref{tab:table2} reports the variables that were included in the final network structure.

```{r table2, results='asis'}
varcutoffs %>% 
  filter(!var %in% c("sdnn", "RR", "I", "GDP", "tlelg", "tlelg_1", "tlelg_2", "tlelg_3", "bal_asym")
) %>% 
  mutate(var = ifelse(var == "hrv", "HRV", var)) %>% 
  kable("latex",
       align = c("llcc"),
       caption = "Variables included in the final Bayesian network structure.",
       col.names = c("Variable", "Definition", "State 1", "State 2")) %>%
  kable_styling(latex_options = "HOLD_position") %>% 
  column_spec(2, width = c("5.2cm")) 
```

@@

During the study, 46% (_n_ = 117) of participants reported at least one injury with an average severity of `r mean_sd(datanet$days_missed, digits = 0)`, days (range = `r min(datanet$days_missed[datanet$days_missed > 1])` - `r max(datanet$days_missed)` days).
Both male and female participants reported a greater number of acute compared to chronic injuries (male, acute = `r inj_mechanism[1,3]`, chronic = `r inj_mechanism[2,3]`; female, acute = `r inj_mechanism[1,2]` chronic = `r inj_mechanism[2,2]`), and non-contact injuries were more common than contact injuries (male, non-contact = `r inj_mechanism[4,3]`, contact = `r inj_mechanism[2,3]`; female, non-contact = `r inj_mechanism[4,2]` contact = `r inj_mechanism[3,2]`).
Table \ref{tab:table3} shows the number and percentage of injury types sustained by both male and female participants.

@@

```{r table3, results='asis'}

@@

The first network structure obtained from the data (Fig \ref{fig:fig1}) examined the interactions between explanatory variables, independent variables and probability of injury across time points in the study. 
Strong sport-related and gender-based connections between several explanatory and independent variables were demonstrated for individual and team-based sports.
The "ind_team" node had strong arcs to hours (`r str_finder("ind_team", "hours")`) and nlebase (`r str_finder("ind_team", "nlebase")`).
Individual athletes were more likely to have "High" hours per week (`r x[5,1]`) compared to team-based athletes (`r x[6,1]`).
Individual athletes were also more likely to have "High" negative life events in the 12 months preceding the start of the study compared to team-based athletes (individual athletes = `r x[7,1]`, team-based athletes = `r x[8,1]`).
The arc from clevel $\rightarrow$ balance_1 had a strength of `r str_finder("clevel", "balance_1")`, with lower level performers more likely to have decreased balance ability (`r x[21,1]`), compared to national level athletes (`r x[22,1]`). 
High gender-based connections were reported for the arcs from gender  $\rightarrow$ stiffness_1 (`r str_finder("gender", "stiffness_1")`) and gender $\rightarrow$ stiffness_2 (`r str_finder("gender", "stiffness_2")`), with males more likely to have "High" stiffness compared to females (males = `r x[14,1]`, females = `r x[13,1]`).
Irrespective of sport or gender, strong connections were found between explanatory variables. The arc from nlebase $\rightarrow$ RI_1 had a strength of `r str_finder("nlebase", "RI_1")`, and the probability of RI_1 being in the “High” state increased from `r x[2,1]` to `r x[3,1]` when nlebase increased from “Low” to “High”.

@@

```{r fig1, fig.align='default', fig.height=10, fig.width=15, fig.cap='The full Bayesian network structure was plotted using the _strength.plot_ function in bnlearn. The strength of each arc is shown graphically by the style of the arc. Thin, dashed arcs indicate the weakest arcs (arc strength below 0.50), whereas thick solid arcs indicate the strongest arcs (arc strength of 1). White nodes in the network indicate the explanatory variables, blue nodes indicate T1_1 and T2_1 variables, and red nodes indicated T2_2 and T3_2 variables. The injured_X nodes have been coloured gold as they are the main nodes of interest within the network.'}
renderGraph(gR)
```

@@

The first conditional probability query (CPQ) investigated the variables in the Markov blanket for injured\_1 (Fig \ref{fig:fig2}), which contained hours spent training per week (hours), negative life events (nlelg_1), muscle stiffness (stiffness_1), current competitive level (clevel) and balance (balance_1). 
The arc between nlelg_1 and injured_1 was fixed in the network, so has the maximum strength of 1.

```{r fig2, fig.cap="Markov blanket of injured\\_1. Arc strengths are included as arc labels.", out.width="100%", fig.align="center"}
include_graphics("figures_doc/Fig2.jpeg")
```

The CPQ for injured_1 in the "injured" state for all variables that were linked to injured_1 is shown in Table \ref{tab:table4}.
The probability of injured_1 = "injured" rose from `r query1[4,2]` to `r query1[4,3]` when stiffness was "High" compared to "Low".
Negative life events had a negligible effect when moving from the "Low" to "High" state.

```{r table4, results='asis'}
query1 %>%
  select(c(1, 2, 3)) %>%
  mutate(variable = str_replace_all(variable, "_", "\\\\_")) %>%
  kable(
    .,
    caption = "Probability of injured\\_1 being in the ``injured'' state, conditional on each variable",
    col.names = c("Variable", "Low", "High"),
    escape = F
  ) %>%
  kable_styling(latex_options = "HOLD_position") %>%
  row_spec(0,bold=TRUE)
```

The second CPQ investigated the outcome of injured_1 being "injured" conditional on all variables in the Markov blanket.
The Markov blanket contained five nodes, each with two possible states resulting in $2^5$ combinations of variables, therefore only the three lowest and highest probabilities are shown in Table \ref{tab:table5} (complete results in supplementary table 4).
The combination of lower competitive level, "High" hours per week, "Low" negative life events, "High" balance and "High" stiffness resulted in a probability of `r max(query2$prob)` for injured_1 being in the "injured" state.
When all variables were in the "Low" state the probability of "injured" was approximately `r min(query2$prob)`

@@

```{r table5, results='asis'}

query2 %>%
  select(7, 2:6) %>%
  rename(Probability = "prob") %>%
  `colnames<-`(str_replace_all(colnames(.), "_", "\\\\_")) %>%
  mutate(clevel = str_replace_all(clevel, "_", "\\\\_")) %>%
  kable(
    .,
    caption = "Highest and lowest probability of injured\\_1 being in the ``injured'' state, conditional on the all variables in the Markov blanket for injured\\_1.",
    escape = F,
    align = "ccccc"
  ) %>%
  kable_styling(latex_options = "HOLD_position") %>%
  pack_rows("Highest", 1, 3) %>%
  pack_rows("Lowest", 4, 6) %>%
  row_spec(0,bold=TRUE)
```

@@

The Markov blanket for injured_2 is shown in Figure \ref{fig:fig3} and contained gender, FFFS_1, stiffness_2, balance_2 and hrv_2 (table \ref{tab:table6}).
The arc between stiffness_2 $\rightarrow$ injured_2 was comparable to the arc between stiffness_1 $\rightarrow$ injured_1.
Very weak arcs (0.3) between injured_2 $\rightarrow$ balance_2 and injured_2 $\rightarrow$ hrv_2 were also present in the Markov blanket for injured_2.
Similar to injured_1, stiffness_2 doubled the probability of injured_2 being "injured" from `r query3[5,2]` in the "Low" state to `r query3[5,3]` in the "High" state.
FFFS_1 in the "Low" state increased probability of injured_2 being "injured" by `r query3[2,2]-query3[2,3]` compared to the "High" state.
"High" negative life events decreased the probability of injury from `r query3[3,2]` to `r query3[3,3]`.

```{r table6, results="asis"}

query3 %>%
  mutate(Low = format(Low, nsmall = 2)) %>% 
  mutate_all(~str_replace_all(., "_", "\\\\_"))%>%
  kable(
    .,
    caption = "Probability of injured\\_2 being in the ``injured'' state, conditional on each variable in the Markov blanket for injured\\_2.",
    col.names = c("Variable", "Low", "High"),
    escape = F,
    align = "cccccc"
  ) %>%
  kable_styling(latex_options = "HOLD_position") %>%
  row_spec(0,bold=TRUE)
```

```{r fig3, fig.cap="Markov blanket for injured\\_2.", out.width="100%", fig.align="center"}
include_graphics("figures_doc/Fig3.jpeg")
```

@@

The three lowest and highest conditional probabilities based on all the variables in injured_2 Markov blanket are presented in Table \ref{tab:table7} (complete results in supplementary table 5).
The combination of "Low" FFFS_1, "High" stiffness_2, "High" balance resulted in the greatest  probability of injured_2 being "injured", with the highest probability of injury being `r max(query4$prob)`.
With all other variables held in the "High" state, the probability of injured_2 being "injured" rose from `r min(fffs_query$prob)` to `r max(fffs_query$prob)` when FFFS_1 was in the "Low" compared to "High" state.
The combination of "Low" stiffness, "Low" balance and "High" FFFS resulted in the lowest probability of injured_2 being "injured".

```{r table7, results="asis" }

query4 %>% 
  toptail(3) %>%
  select(7,2:6) %>%
  rename(Probability = "prob") %>%
  `colnames<-`(str_replace_all(colnames(.), "_", "\\\\_")) %>%
  mutate_all(~str_replace_all(., "_", "\\\\_"))%>%
  kable(
    .,
    caption = "Highest and lowest probability of injured\\_2 being in the ``injured'' state, conditional on the all variables in the Markov blanket for injured\\_2.",
    escape = F,
    align = "cccccc"
  ) %>%
  kable_styling(latex_options = "HOLD_position") %>%
  pack_rows("Highest", 1, 3) %>%
  pack_rows("Lowest", 4, 6) %>%
  row_spec(0,bold=TRUE)
```

@@

The second network structure (Fig \ref{fig:fig4}) examined changes within variables between time points and the probability of injury/
An arc between BIS $\rightarrow$ FFFS with strength `r str_cfinder("BIS", "FFFS")` was present in the network.
Arcs between clevel $\rightarrow$ BIS and gender $\rightarrow$ stiffness had a strengths of `r str_cfinder("clevel", "BIS")` and `r str_cfinder("gender", "stiffness")` respectively
Similarly, the arc between hrv $\rightarrow$ FFFS was `r str_cfinder("hrv","FFFS")`.
The arcs between BIS $\rightarrow$ FFFS and hrv $\rightarrow$ FFFS were examined further by drawing random observations from the conditional probability distribution and examining the relationship in a Bayesian linear regression model.
A separate linear regression examined the interaction between BIS and hrv to be examined.

```{r fig4, fig.cap="Network structure of the changes within variables between time points.", out.width="100%", fig.align="center"}

include_graphics("figures_doc/Fig4.jpeg")

@@

Results from the Bayesian linear regression model are presented in Table \ref{tab:table8} and include 95% credible intervals (CrI).
Increases in BIS were associated with increases in FFFS ($b$ = `r bmtable[3,2]`, 95% CrI = `r bmtable[3,4]`), whereas positive changes in hrv where associated with decreased changes in FFFS ($b$ = `r bmtable[2,2]`, 95% CrI = `r bmtable[2,4]`).
There was no clear interaction between hrv and BIS ($b$ = `r bmtable[4,2]`, 95% CrI = `r bmtable[4,4]`).

```{r table8, results="asis"}
bmtable %>%
  kable(
    .,
    caption = "Estimate, error and 95\\% credible intervals for the fixed effects in the linear model containing FFFS, BIS and hrv.",
    col.names = c("Term", "Estimate", "Error", "95\\% CI"),
    escape = F,
    align = "lccc"
  ) %>%
  kable_styling(latex_options = "HOLD_position") %>%
  row_spec(0,bold=TRUE)

@@

The Markov blanket for the "injured" node contained previous injury, gender, training hours per week, stiffness and nlec (Fig \ref{fig:fig5}).
For stiffness and nlec the values in the nodes represent the standardised change between time point.
Combinations of nlec and stiffness at one SD below the mean change, at the mean change, and 1 SD above the mean change are presented in table \ref{tab:table9}.
Increases in muscle stiffness was found to increase the risk of injury, which was further increased when there were increases in NLE stress. 
Changes in both nlec and stiffness of 1SD above the mean change resulted in a high probability of being injured (`r query6a[9,1]`) over the duration of the study.
With stiffness held at the mean change, the probability of "injured" rose notably from `r query6a[4,1]` to `r query6a[6,1]` with nlec at 1 SD below an 1 SD above respectively.

```{r fig5, fig.cap="Markov blanket for the injured node in the network reflecting changes within variables between time points.", out.width="100%", fig.align="center"}
include_graphics("figures_doc/Fig5.jpeg")
```
